%!TEX root = ../Master_Thesis.tex

% Wie schreibe ich das Abstract?
% Bei der Vorbereitung des Abstracts sollten Sie sich an folgenden grundlegenden Punkten orientieren:

%% Motivation des Textes:
%% worin liegt die Bedeutung der entsprechenden Forschung, warum sollte der längere Text gelesen werden?

%% Fragestellung:
%% welche Fragestellung(en) versucht der Text zu beantworten, was ist der Umfang der Forschung, was sind die zentralen Argumente und Behauptungen?

%% Methodologie:
%% welche Methoden/Zugänge nutzt der Autor/die Autorin, auf welche empirische Basis stützt sich der Text?

%% Ergebnisse:
%% zu welchen Ergebnissen kam die Forschung, was sind die zentralen Schlussfolgerungen des Textes?

%% Implikationen:
%% welche Schlussfolgerungen ergeben sich aus dem Text für die Forschung, was fügt der Text unserem Wissen über das Thema hinzu?
 
\begin{abstract}

This thesis deals with a configuration-based application infrastructure, which allows cross-platform mobile application development for iOS and Android.
The key features are that non-developers can create app configurations and that updates can be loaded automatically over-the-air without using App Store or Google Play Store.
The topic is introduced by outlining use cases and portraying example applications.
The capabilities that arise from using configurations are exposed by applying model theory and model-driven development techniques to the interpreter and comparing the used approach to model-driven architecture.

Since there is an existing and productive implementation, the thesis will evaluate requirements and the architectural quality.
An alternative implementation approach is introduced, which utilizes abstraction and platform independence.
The two options are compared by using a value-benefit analysis and additional aspects like continuous reengineering and hybrid approaches are mentioned.

The evaluation shows, that the alternative and abstract implementation approach implements the requirements at a higher quality.
But the difference is only marginal, so no clear statement can be derived from it.



\end{abstract}
\cleardoublepage