\chapter{Introduction}

Mobile application and digital media usage is still on the rising\cite{perkins_internet_2018}.
Apps are one of many possible ways to interact with customers and consumers.
A web appearance on social media or a dedicated web page can also fulfill these needs.
The way, the user interacts with a web page is more shallow, because of lower expectations.
A user needs to log in most of the time.

A mobile application can enable user specific features without logging in, because the user can be identified with an identification token, that is generated, while the user installs the app.
It is possible to activate special features like geofencing, push notifications or local data storage, that is synchronized with the platform cloud services (e.g. Apple iCloud or Google Drive).

A mobile app integrated more deeply into the device, than a web app.
However, web browser already support a lot of the features by now, that originally were only available to native apps.
The current development indicates, that the differences will be further eliminated by the browser vendors.

\section{Problem definition}

Mobile applications have significant disadvantages, that push the development of web apps forward.
Changes have to be approved by the phone vendors, in order to be able to release updates to the app.
This causes development delays and reduces the ability to react to events or customer requirements.

In addition, managing a lot of apps (> 20) requires a lot of work, since all of them have to be maintained.
New operating system features have to be implemented in all of the apps, which can cause inconsistencies.
All apps have to be tested in detail, to make sure that all apps implement the required change correctly.

Developers need deep system knowledge to develop the apps.
Every change has to go through one of each platform experts, so development is expensive.

\section{Motivation}

To be able to deliver the full mobile application experience, the app has to be native.
This will change, when all of the phone's hardware interface is exposed to the web browser.
But for now, the only way to use these features and be good on battery life and data usage is by providing a native app, that is written in the native language and uses features, directly provided by the phone vendors.

Web app wrapper like ''ionic'' are able to use these features and provide their own interface to the web app.
In this way, a web app can in fact use restricted components like push notifications and beacons.

With this approach, every new app still has to be developed and old ones have to be maintained individually.
In addition, it involves the dependence on third party framework vendors. 
Building apps with ionic creates a massive list of dependencies: ionic, angular, npm and nodejs.
Native apps have the ability to reduce reliance, by only using native libraries, which are very comprehensive by default.

To be able to create many apps with little effort, a model driven approach can be considered.
A well abstracted model can have the ability to be easy to understand by users and the platforms can bring their respected implementation in, so any app looks and feels native.
A service can deliver pre-rendered sub-models, so that complex client workload can be moved to the server.

Existing apps are easy to maintain, because there is only one code base per framework.
One bug fix applies to all apps.
Also, it is easy for anyone to change small things like a label, a background color or the sorting of a list.
The goal is, to be able to configure new apps in one day, together with the customer.


\section{Distinction}

This thesis will not describe the development experience nor the initial expense of the project. Neither will this thesis evaluate financial aspects and at which point, the project was or is worth the development costs.


\section{Structure}

\subsection{Research Question}

The thesis aims to answer the question ”Can a code interpreter be integrated in any domain?”. In addition, the development approach is compared to procedu- res, that are described in popular literature.

kevin rulez