% Alles was man wissen muss um das Konzept zu verstehen.

\chapter{Background}

The system, that is discussed here, is a model interpreter that is designed to run mobile applications on iOS and Android.
The system, which includes the interpreters and the backends are called \ac{mos}.
It is in active development and poweres over 50 apps that all are available in the Apple AppStore and Google PlayStore.
The app categories range from retail to event to city marketing apps\footnote{\url{https://fabrik19.de/technologie/}}.


\section{Company}

The creator of the system is Fabrik 19 GmbH.
The company is based in Gießen, Hesse, Germany and was founded 20XX by Mark Pralle and Sven Herchenhein.
At the moment, the company has bout 25 employees.
Besides of apps, there are also other projects and solutions, like consulting and webhosting.

\subsection{Markets}

Fabrik19 does not deliver apps directly to customers, but is partner of retail corporations and city marketing agencies.
The current flagship is the Gießen App, which was developed in cooperation with DISTAMA GmbH.
The app was downloaded 34,000 times and has 20 distinct and active users per day.

Besides mobile apps, digital signage is another important market.
The companion to the Gießen App is a stele that contains a large touchscreen and runs an app that is optimized for big touchscreens.
It is located on the main pedestrian shopping street in Gießen and provides respective informations on the stores and points of interests\footnote{\url{https://distama.de/bid-seltersweg/}}.
Other stele systems are also active in Globus hardware stores and Globus grocery stores.

\subsection{City Marketing}

Since the apps use live data and have the ability to change it's content structure, an active support and maintenance is needed to keep the app alive.

To be able to advertise and publish city marketing apps as a service, Distama GmbH was founded which is a subsidiary firm of Fabrik19.
The goal was to combine technology knowledge with marketing, to be able to provide the app and the attached customer support.
City marketing agencies are not always able to support an app, because they lack experience and domain knowledge.
Distama is the mediator between the developer, the customer and the user.



\section{Model Interpreter System}

The main development focus lays on the app interpreter platform ''mos''.
It supports the two major platforms iOS and Android, but there is also a lightweight version available in .Net, which is running on the steles as a Windows app.
A web client is also actively in development, which aims to replace the .Net version.

A configuration are multiple JSON files, that represent the app meta information, its layout and its behavior.
The files are called node.
There is one root node, whose url is baked into the app while compiling.
This root node defines the first screen of the app.
All the other screens can be defined in other nodes.
When displaying another node, the new node is loaded from cache or network and its new view is pushed on the view stack, which is done by the interpreter.
On the particular platforms, this action will look different and will use the platforms native animations.
Web apps like Ionic have their own push animations which makes them feel foreign.
The previous node can pass data, like the ID of the item that should be displayed.

The goal is, to be able to create a new app in one day, by meeting with the customer and configuring the app together.
The syntax of the configuration is designed to abstract all of an apps standard behavior.
Because of a custom layout system, that can be described with constraints and will layout automatically, it is not possible to recognize that it is only a configured app, instead of a custom native one.

The apps have sophisticated caching and optimization mechanisms, which makes it possible to use them completely offline, without any additional configuration effort.
The configuration, images and data are cached heavily, which makes the apps very fast.
But because of the dynamic design of the configuration syntax, it is possible to clear the caches and force new versions.
The cache first policy of the interpreter makes sure, that contents are always available and loads changes in the background.

Besides of displaying data, the client is highly intractable and can perform various custom tasks on the device, like uploading local data, saving data to the devices user settings and providing push, beacon and location infrastructure.

This data can be used in describing the layout by defining layer conditions, which makes the app behave accordingly.


\subsection{Use Cases}

The main use cases are marketing and customer interaction.
The concept is data source driven and enables configuration of any data source.
Because of a template engine, data can be arranged for display in an intuitive way.

Since the configuration can be changed on the fly and without the need of providing an update via app store, the contents can be adjusted to fit current events like holidays or promotions.
Especially for digital signage it is important to be able to change the contents without any application updates.

Any change to the configuration is reflected as soon as the app checks for configuration changes.
The update mechanism are also highly customizable, so that the optimal update intervals and strategies can be used for different apps.

A city marketing app might need a medium update interval, so that promotions and special content are available when needed.
A event app should have a very short update interval, so that the events program changes or announcements are delivered as fast as possible.


\subsection{Interoperability}

The app can be filled with data via common REST services, that deliver JSON data.
The app uses HTTP mechanisms and a custom delta system to only load new data.
The mos backend system also has a data proxy, which can load any data and provide it in a simpler, preprocessed format, so it is simple for the apps to download.
This way, it is possible to use any data source and still be able to change the templates and layouts without updates.

The apps have the capability to be integrated in a Wordpress instance, a chat service like Mattermost or an secured API that uses authentication methods like client certificates, auth headers or JSON web tokens.

By default any network communication is secured with an app specific client certificate.
Other authentication methods have to be configured.

\section{Models}

Since the mos-Framework tries to abstract mobile application platforms into a common language, the configuration could be considered to be a abstract model of a mobile application.
To classify and evaluate the model approach for this domain, the following section will discuss models and their properties.

\subsection{General Models Specification}

A general model is described by Herbert Stachowiak\footnote{\cite{stachowiak_allgemeine_1973} 131}.
He defines three properties, that apply to all models.
A model is a representation of an original, that is either a natural or artificial thing, that can have properties and behavior.
The original can be a model it's self.

The second property is the ability to reduce properties and behavior of the original to a set, that is relevant for the model handlers.
The original might be a complex system and only a subsystem or specific properties and their interaction are relevant.
For this, the model can reduce the complexity, to be more simple to understand and to minimize the error rate.

Models can not be clearly assigned to their originals.
That means, that original and model are not equal, but can be interchanged.
There are three cases defined by Stachowiak, in which the model can be used as an replacement.
Models are not only models of something, but they are also models for someone, a person or an artificial model user.
They perform their functions within a time interval.
And they are models for a specific purpose.
Often times, the purpose is an action that is not possible on the original because it may be to complex, or an physical original could be destroyed.

If theses properties apply an object, i can be called model according to Stachowiak.
It can be applied by numerous of common usages of the word.
Mathematical models, philosophical models, simulations and concrete object models are some examples, that fit the requirements.

\subsection{Descriptive Models}

\subsection{Prescriptive Models}

\section{Model Driven Development}

\subsection {Code Generators}

\subsection {Model Interpreters}

