%!TEX root = ../Master_Thesis.tex

\chapter{Concept}

This chapter will describe the goals of \ac{mos} and it's iOS implementation.
The first section will describe the informal requirements to the model.
Also, the model's evolution and progression is discussed, which is relevant to the main idea behind the model interpreter.

\section{Implementation Goals}

The main problem of mobile app development is the iteration process, which consists of collecting requirements, creating conceptual prototypes and implementing on the two main platforms.
Since requirements are informal, incomprehensive or change during the development, multiple iterations have to be done to converge to the final product.

To minimize implementation times and be able to do more iterations, \textit{Same Day Delivery}\footnote{The term originates from parcel shipping and means that a parcel arrives at the customer on the day of the order.} tries to eliminate the times, that is independent to the customer.
That means together with the customer the app is designed and the model is created in one session.
Since the functional scope is already defined and all of the feature are implemented in the interpreter, the domain expert who is directing the meeting can create the model live and on the fly.
The App can be tested by the customer directly, without the need to compile an custom app.
Common tasks like downloading data from a web source, forms and user interaction are configured on a high level, so the customer is able to understand aspects of the app.

\subsection{Requirements}

To be able to validate the meta model and an implementation, requirements have to be defined.
A concrete model also has requirements, which are application specific however.
That's why, only the meta model and the implementation are considered.
Since this chapter is about the concept of the \ac{mos} interpreter, the discussion on the requirements of the implementation are found in chapter \ref{chap:implementation}.

The following paragraph lists the functional requirements to the meta model.
Those are aspects that have significantly influenced design choices.

\begin{enumerate}
	\item \textbf{Separation into independent sub-models}
	
	An app loads a root node, which then can navigate to a different node, or show it's preview.
	Every node is theoretically independent and can be used as root node.
	To enable interaction between nodes, an action that displays a new node can inject additional data into the node.
	An example is a master-detail-view, which is a common pattern in application design.
	A master view is a list of all objects.
	It knows everything and is connected to the database.
	When the user clicks on one of the objects, the detail view is presented by passing either the database ID or the whole object as parameter.
	The same approach is possible with nodes, which results in the detail view node not being independent to the master view node.
	
	\item \textbf{Fully customizable layouts}
	
	A layout can be defined by arranging different kinds of layers by configuring layout constraints. Also, native navigation bar, tab bars and side menus can be set independent to the layout.
	All of the common layer types are supported like text, image and color.
	Those layers have layer parameters that make them customizable.
	Position and size can be defined with constraints, by binding the edges of a layer to the edges of other layers or the superior container.
	The binding can be modified with additional margins so that dynamic and portable layouts can be defined.
	The system is inspired by Apple's Auto Layout\footnote{\url{https://developer.apple.com/library/archive/documentation/UserExperience/Conceptual/AutolayoutPG/index.html}}.
	
	\item \textbf{Reactive behavior}
	
	Everything that is depended on something else needs to update on changes.
	For example: If a layer is depended on an \ac{SQL} query and the corresponding tables change their contents, the layer needs to update itself with the new data.
	Another example are layers that have a form data key and thereupon synchronize their contents with the view's form.
	Layers can also have a setting data key, which enables a layer to synchronize it's value with a field in the database.
	This makes it possible to use the value in \ac{SQL} queries, apply it to templates or uploading data to a service.
	
	An example that enables a React\footnote{Reactive and functional UI framework. \url{https://reactjs.org/}}-like behavior is the \textit{layout layer}.
	It's parameters are an \ac{SQL} query, a layout name and a template name.
	The layer runs the query and applies the data to the template, and the template to the layout and displays the layout.
	If one of the tables that are selected in the query changes, the \textit{layout layer} re-renders itself.
	
	\item \textbf{Separation of data, templates and layouts}
	
	A template defines layer parameters by taking a dictionary of data and applies it to itself. The result can then be applied to a layout.
	
	\item \textbf{Database definitions}
	
	An \ac{SQL} database can be designed, that is in sync with a data source.
	
	\item \textbf{Actions}
	
	An action can perform common tasks like uploading data, setting layer values, or triggering fetches. Action can be fired by clicking on a layer or are triggered by other actions, which take the result of the previous action as context data to use.
\end{enumerate}

The following paragraph lists the non-functional requirements.

\begin{enumerate}
	\item \textbf{Easy to understand}
	
	Existing models are easy to understand. Names are clear and there is no room for interpretation.
	
	\item \textbf{No boilerplate code}
	
	All of the configuration contains highly concentrated information and no boiler plate code has to be defined.
	
	\item \textbf{Easy to parse by an interpreter}
	
	The model is easy to parse and is encoded in a common format.
	Validators can check for simple errors.
	
	\item \textbf{Backward compatibility}
	
	Adding new functionality has to be added without any old interpreters to fail on new models and a new interpreter to fail on old models.
\end{enumerate}

All of the above functional and non-functional requirements are developed both with an bottom-up and a top-down approach.
Since the interpreter contains most of the business logic and implicitly defines the meta model, it is one of the key aspects.
That is why, the meta model is structured to be parsed and interpreted in a simple and direct way.
In the same way, the usability of the models are also considered, so that both sides are balanced.


\section{Model Syntax}

To meet the majority of the non-functional requirements, \ac{JSON} is used to define the model.
It allows the client to parse common structures and define optional and default values.

The model is divided into sub-models.
The first model is called root node and is loaded by the interpreter on the first app start.
The \ac{URL} to the root node is compiled into the app.
A node defines some of the meta data, static data, a starting layout reference and events.
The meta data contains an \ac{URL} that is used by the client to synchronize layouts, templates and database definitions.
Database definitions also contain \acp{URL} that are requested if needed to synchronize actual content data.

The client implementation is able to optimize the data usage by only reloading data that has changed.
\ac{HTTP} techniques like ''etag''\footnote{Entity Tag, a unique identifier assigned by the web server to specify a resource and it's version.}, ''modified-since''\footnote{A \ac{HTTP} header containing a timestamp.} or delta updates\footnote{Requesting a resource by \textit{/api/file.60.json} will deliver the changes of the last 60 seconds.} are used to optimize the update process.
The model's job is to interconnect web services and the client.
Optimizations can be implemented without changing the model, since they are not coded into the model in any way.


\subsection{Drawbacks}

boundary conditions

limits to development of app contents (app is out, but the customer needs more functionality, which is not possible: incomprehensible reasons for limits.)

\subsection{Iterative Development}





















