%!TEX root = ../Master_Thesis.tex

\chapter{Implementation}
\label{chap:implementation}

In this chapter, the approaches that are described in chapter \ref{chap:concept} are demonstrated using code examples and the real reference implementation.
In the first section, the initial state of the interpreter is explained.
This is the starting point from which the re-engineering options are applied.

In the following section, the first approach is examined.
This is the option that is described in section \ref{sec:nativeapproach}.
The subsequent section will outline a possible virtual machine implementation exemplary.

\section{Initial State}
\label{sec:initialstate}

The original iOS interpreter is written in Objective-C and uses \textit{AsyncDisplayKit}\footnote{AsyncDisplayKit was renamed to Texture. \url{https://github.com/texturegroup/texture/}} as an asynchronous \ac{UI} framework.
It allows instantiation and lay-outing of view components on any thread.
This means, that load intensive tasks can be swapped to background threads.

This makes synchronization of \ac{mos} specific behavior like condition evaluation and state management harder.
Also, custom view implementation that are used very often are more challenging to build, because they have to be a subclass of \textit{ASDisplayNode}, which represents one simple view component.
In addition, the framework's future was unknown, since it has not been maintained and was licensed by Facebook.
It has no native Swift \ac{API}.
Later, the framework moved to Pinterest, where it has been maintained more regular.

Because of the high dependence to \textit{AsyncDisplayKit} and the difficult maintenance, the next iteration of \ac{mos}, which should use Swift, needs to move away from the \ac{UI} framework.
The code basis that was used to utilize the framework became obsolete because native views, that are enriched with \ac{mos} behavior, should be used.

\section{Native Approach Implementation}

Instead of \textit{AsyncDisplayKit}, that is described in \ref{sec:initialstate}, a approach should be used, that primarily uses the native \ac{UI} and the main thread.
The main approaches will be outlined and explained in the following sub sections.

\subsection{Native User Interface}

Every layer operates directly on their native representation.
That means that if changes need to be applied, the layer manipulates it's \ac{UI} component directly.
This is necessary because some of the layouting calculations are based on native \ac{UI} \acp{API}.

An example would be the size of text in a label layer.
A text layer can be configured to update it's size automatically based on the text content.
Also, a maximum line count could be defined.
Since the native text \ac{API} only supports the calculation of the text size based on a constrained size, but not based on a line count, another approach must be taken.
In this case, the text and the calculated text size is applied to the label.
After that, \textit{sizeToFit()} is called on the layer.
It's a native method, that shrinks the label based on the maximum line count.
The resulting size is then used as the layer's frame to layout the remaining layers.

Another example is the size of all sub view of a grid layer.
Since a native grid layouts views based on a collection layout definition, the resulting size can only be determined by asking the grid on the main thread.

This means, that lay-outing and conditions happen directly on the \ac{UI}.
It keeps the code client and simple configurations are run fast.
The following listing outlines a layer and it's behavior.

\begin{lstlisting}[style=swift, caption={Layer class (excerpt)}, label={lst:layerclass}]
class Layer {
	let baseLayerParameters: BaseLayerParameters
	var state: LayerState
	weak var superlayer: SublayeredLayer?
	
	init(baseLayerParameters: BaseLayerParameters)
	func performActions()
	func applyConditions(afterContextChanges contextChanges: [ContextChange] = []) -> [ContextChange]
	func apply(databaseChanges: [DatabaseChange]) -> [ContextChange]
	func setNeedsLayout()
	func size(constrainedBy constrainedSize: CGSize) -> CGSize?
	func activate(state: LayerState) -> LayerStateChangeSeverity?
	func view() -> UIView
	func update(view: UIView)
}
\end{lstlisting}

Listing \ref{lst:layerclass} outlines an excerpt of the Layer class.
The function \textit{update(view:)} is the bottleneck with the most significant impact.
It's responsible for relayouting it's native components.


\subsection{Concurrency}

Since changes are happening directly on the \ac{UI} elements


\section{Virtual Machine Implementation}




















