%!TEX root = ../Master_Thesis.tex

\chapter{Evaluation}

This chapter will take the implemented interpreter and the meta model of an app configuration and will classify the development approach.
The properties of general models as described in section \ref{generalmodeltheory} will be compared to the meta model definition of \ac{mos}.
The comparison also includes the development approach of the meta model.
The meta model and its instances will be examined for general validity and inconsistencies.

In the next section, the implementations of the interpreter are evaluated.
This will be the focus, since the interpreter is the main subject in this thesis.
The requirements as described in section \ref{sec:implementationgloas} will be weighted by individual comparison.
The old and new implementations will then be rated based on their compliance with the requirements.

The evaluation is needed to be able to compare the two implementations and ultimately decide, if the new implementation is a step forward.
Also, there is the question if it is needed to re-engineer the interpreter for a third time, based on the \ac{VM} approach of section \ref{sec:virtualmachneimplementation}.
Because the first approach was not perceived as a much better or faster client, a totally new approach has been considered, which is the \ac{VM} structure.

The answers to the research question of section \ref{sec:goals} will be backed by the result of the evaluation.
This will then lead to the final conclusion.


\section {Meta Model}

This section will examine the meta model and will compare the approach of developing and evolving with common, established procedures.
The meta model development is independent to its later runtime.
A meta model development for an interpreter can be compared to the development of meta models for code generation, because they represent the same original.
This means, that the development approach of section \ref{sec:modelingasoftwaresystem} can be applied.
But since models reduce their complexity for a specific use case, they can only be compared in regard of a common use case.
In this evaluation, the use case is a model interpreter.

\subsection{Development Approach}

Since a model interpreter is not code centric, but focuses on features and requirements, the development approach of separating the code of an existing original (\ref{sec:modelingasoftwaresystem}) differs slightly to the one used for the development of the \ac{mos} meta model.

The approach that was described in figure \ref{fig:mdsd1} begins by analyzing the application code of existing originals.
Instead of analyzing source code, the requirements of generic mobile applications have been analyzed and divided into the three aspects.
This was done, because no direct originals existed.
Also, it is not mandatory to develop an interpreter based on original code.
It is more practical to base an interpreter implementation on original abstract behavior.

\textit{Schematic repetitive code} refers to code, that is abstracted into the model and later generated by a code generator.
This can be applied to the model interpreter because it refers to the repetitive behavior of a mobile applications.
It contains all the features that the application offers, that is described by the model.
Some code could have been taken directly from an original mobile application, because some design patterns of mobile applications also require a strong abstraction, so the \ac{UI} code is generic anyway.
However, most parts of the interpreter code is not found in application code, because it implements the abstract and very specific mos behavior of nodes, events and actions.

\textit{Generic code} is domain specific code, that is used in all apps and is not part of the model.
Since mobile application frameworks are already abstracting a lot of work, most of the code that is used in the interpreter is directly relevant for model interpretation features.
If there would be no native \ac{UI} framework and every app must implement it's own, this would be part of generic code.

\textit{Individual code} can be mapped to interpreter features that have been developed for single specific applications.
An example for that are tool layers, that display a ruler, a goniometer and a mechanic's level.
Those layers have only been developed for a single app which implements hardware store online services.
This might not be custom code, that has been injected, but it is a model extension that was developed instead of providing a code interface.
Custom code in the domain of platform independent applications is not goal-oriented, because the main idea is to reduce the need to write duplicated code in different languages.
Also, a interface for dynamic native code extensions increases the complexity of app generation.



\subsection{Structure}

The meta models structure was directly designed to be interpreted on mobile devices with fluctuating internet connectivity.
That's why, all the components can be loaded separately and cached.
Nodes, which define the basic view structure of an app, can be loaded lazily and on demand.
The implicit behavior of the interpreter supports the models intentions, by operating cache first and loading only if the various hash values changed.
This is done by calling the \ac{API}'s update interface which provides hashes for the different elements.
For the model, this means that it's performance is directly depended on the providing web services.
Therefore, the model is only a small part of a large infrastructure.

To evaluate the meta models platform dependence, all the features, that are directly referencing mobile devices will be outlined.
It is possible to define a nodes style based on the phone's device orientation and screen size.
For screen size, the identifiers \textit{phone}, \textit{tab7} and \textit{tab10} are available, which represent a normal phone, a seven inch tablet or a ten inch tablet.
The device orientation can be distinguished by the values \textit{all}, \textit{portrait}, \textit{landscape}, \textit{portraitsensor}, \textit{landscapesensor}.
Those configuration options are valid for any mobile device platform like tablets and phones.
But it turned out that it can be also used directly for interactive digital signage devices, because they behave like a large tablet using their touch-based interface.
Android and iOS devices are perfectly capable of mapping those identifiers to native values.
A computer running desktop systems like Windows cannot directly provide those values, but instead can simulate a mobile device environment by providing \textit{landscape} and \textit{tab10}.
There are no further ways to distinguish the main platforms iOS and Android, since mos is designed to only support features that cover both platforms.

\subsection{MDA Abstraction}

The abstraction is a key element of model-driven development and is directly relevant to the model syntax.
The \ac{MDA} standard defines abstraction levels as described in section \ref{modeldrivenarchitecture}.
Most aspects of the model syntax that is used in mos are computation independent.
An example are the layout constraints that define rules of layouting layers relatively to each other.
Since there are multiple algorithms that can evaluate the rules to concrete frames, this aspect is considered computation independent.
Another example is the declarative synchronization of layer data using form data keys.
Layers with the same form data key share their data.
Since this can be implemented by any platform in any way, this aspect is also computation independent.

There are other aspects that have to be discussed, because they cannot be identified as computation independent.
An app can load data into databases that can be configured using the model.
This configuration can be interpreted by any relational database system, since only commonly used column types can be used.
This can be considered computation independent, because the internal structure and processing of the configuration is not defined.
But since concrete \ac{SQL} queries can be used in the definition, this part of the model cannot be considered computation independent, since there is a direct and formal definition of the data syntax and processing.
Instead, this case of \ac{SQL} usage can be considered platform independent.



\subsection{Requirements}

The meta model requirements of section \ref{functionalmodelrequirements} will now be weighted.
This is done using pairwise comparison.
Figure \ref{fig:pairwisecomparisonmetamodell} shows the intermediate score values, while table \ref{table:pairwisecomparisonmetamodelresult} lists the results.

\begin{table}[h]
	\begin{center}
		\begin{tabular}{l|l}
			Requirement                               & Weight (Score) \\ \hline
			F6 Actions                                   & 9     \\
			F5 Database definitions                      & 5     \\
			F2 Fully customizable layouts                & 5     \\
			F1 Separation into independent sub-models    & 3     \\
			F3 Declarative and reactive behavior         & 1     \\
			F4 Separation of data, templates and layouts & 1     \\
			NF2 No boilerplate code                       & -3    \\
			NF4 Backward and future compatibility         & -5    \\
			NF1 Easy to understand                        & -7    \\
			NF3 Easy to parse by an interpreter           & -9   
			\\ \hline
		\end{tabular}
	\end{center}
	\caption{Result of pairwise comparison of the meta model's requirements.}
	\label{table:pairwisecomparisonmetamodelresult}
\end{table}

Every requirement was compared to each other and a score of -1 or 1 are assigned.
There are ten requirements, so the maximum score is 9, while the minimum score is -9.
The scoring shows, that F6 Actions is the most important requirement.
It is notable, that actions have not been the first thing, that the model offered.
The meta model started with F1 Separation into independent sub-models.
But today, the actions are one of the most important features, while there are workarounds for the other requirements.
Another notable finding is, that the non-functional requirements make up the entire bottom half of the requirements table.
This is, because non-functional requirements on a meta model only affect application and infrastructure developers and do not make any statement about the quality of the final application.

The current meta model, that is used in the various production apps is fully implementing all of the acceptance criteria.
Those are described in section \ref{functionalmodelrequirements} and the following.

\section{Architecture Analysis}

The analysis of a software architecture is described in \cite{klein_evaluating_2001}.
It is one aspect of a software development project and is done before the first design planning of the software's architecture is created.
The architecture of a software is a combination of multiple points of view and abstraction levels.
A software architecture and it's analysis is "a bet, a wager on the success of a system" (\cite{klein_evaluating_2001} 19).
This means, that an analysis might steer the software development process in a possibly reasonable direction, but the actual outcome relating to the requirements is undefined.
In addition, the authors suggest, that an unsuitable architecture can cause a complete project failure.
Also, the software architecture is directly related to the project structure and the team allocation.

An architecture analysis starts with the definition of quality goals, which are non-functional requirements on the system.
Those are not only end-user-oriented, but also include development and maintenance quality goals.
Those goals have to be defined precisely since statements like "The system must be stable." do not express any practical demands.
To make the statement more meaningful, it must name a concrete scenario in which the system can handle errors and is able to stay available.
The following list states some examples.

\begin{enumerate}
	\item The system must stay available after 10.000 simultaneous login requests, where 10\% are unsuccessful logins.
	\item The system must be available ten minutes after a power outage.
	\item The system must have a maintenance mode, in which major components can be swapped, while the user sees a maintenance view.
\end{enumerate}

These goals now imply consequences.
A monolithic architecture might allow a simple and fast development process and fulfills the second goal, but it is not possible to change components while there is still a maintenance view for the end user.
Also, depending on the complexity of a login request, the monolithic approach might be unable to handle those login requests.
A distributed system might be more complex, especially the second goal can be challenging.
But now it should be possible to handle many requests and implement a maintenance mode, where sub systems can be shut down and swapped, while the main handler shows the maintenance view.

A suitable architecture meets it's quality goals and is actually buildable.
This is why the largest part of an architecture evaluation is the specification and prioritizing of quality goals.
The goals are always referencing specific scenarios or requirements.

\section{Requirement Analysis}

The following sections will evaluate the interpreter and the requirement implementation.
The evaluation approach is a slightly modified version of a value-benefit analysis\footnote{This term is not well defined and combines multiple similar approaches. What is actually used here is the Nutzwertanalyse, which became known through \cite{zangemeister_nutzwertanalyse_1976}.}.
The requirements of the interpreter are weighted by their importance.
A detailed documentation is shown in figure \ref{fig:pairwisecomparisoninterpreter} and each of the requirements are described in section \ref{functionalinterpreterrequirements}.
Every requirement is scored by using a pairwise comparison.
The requirement that is more important gets a sore of 1 while the other one is scored -1.
The results are shown in table \ref{table:pairwisecomparisoninterpreterresult}.

\begin{table}[h]
	\begin{center}
		\begin{tabular}{r|l|l}
			& Requirement                               & Weight \\ \hline
			NF1 & Low battery usage                                   & 6     \\
			F5 & Poor configurations lead to fall-backs                     & 6     \\
			F1 & Reflects changes to the model                & 6     \\
			NF2 & Fast app start    & 4     \\
			NF3 & High UI and animation performance         & 4    \\
			NF5 & Small download size    & 2     \\
			F2 & Offline capable                      & -2    \\
			F3 & Cache-first policy        & -6    \\
			NF4 & Low network usage             & -6    \\
			NF6 & Small app size on disk       & -6   \\
			F4 & Clear cache and data       & -8   
			\\ \hline
		\end{tabular}
	\end{center}
	\caption{Result of pairwise comparison of the interpreter's requirements.}
	\label{table:pairwisecomparisoninterpreterresult}
\end{table}

There are 11 requirements, so a maximum score of 10 or a minimum of -10 is possible.
The fact that the scores do not peak at the maximum and the values are more spread out show, that the weights are more similar to each other than the ones of the meta model.
This is, because the requirements are all defined as equally important and the weighting is mostly based on user experience.

\subsection{Native Approach}

The native approach is the currently used production architecture, that is powering most of the company's mobile applications for iOS.
It is comparable to a monolithic approach, since everything is happening on the main thread, and mos specific functionality is directly mapped to native \ac{UI} elements and frameworks.
One example is the evaluation of layout constraints.
If a text layer has a variable height, it should fit it's height to the size of the current text content.
This is calculated using a native iOS Framework, which must operate on the main thread.
Also, layouting the layers multiple times, because of resulting changes of conditions, is directly performed on the native \ac{UI} elements.
This makes the main thread very busy, so the responsiveness and animation quality is not always guaranteed.
The other goals are all easily fulfilled, using the native approach architecture, because they are common mobile application requirements, that are already implemented in the native frameworks.

The next section will describe how well the native approach interpreter is implementing the requirements.
This is the implementation quality.
A scale of 1 to 4 is used:

\begin{itemize}
	\item \textbf{1} This requirements is not implemented
	\item \textbf{2} This requirement is implemented poorly
	\item \textbf{3} This requirement is partially implemented
	\item \textbf{4} This requirement is fully implemented
\end{itemize}

Table \ref{table:implementationscorenativeapproach} will score the requirement based on the implementation quality.


\begin{table}[h]
	\begin{center}
		\begin{tabular}{r|l|l}
			& Requirement                               & Quality \\ \hline
			F1 & Reflects changes to the model                & 4     \\
			NF4 & Low network usage             & 4    \\
			NF2 & Fast app start    & 4     \\
			NF1 & Low battery usage                                   & 3     \\
			F5 & Poor configurations lead to fall-backs                     & 3     \\
			NF5 & Small download size    & 3     \\
			F2 & Offline capable                      & 3    \\
			F3 & Cache-first policy        & 3    \\
			NF6 & Small app size on disk       & 3   \\
			F4 & Clear cache and data       & 3      \\
			NF3 & High UI and animation performance         & 2 
			\\ \hline
		\end{tabular}
	\end{center}
	\caption{Implementation score of each requirement of the native approach interpreter.}
	\label{table:implementationscorenativeapproach}
\end{table}

The requirement weights of table \ref{table:pairwisecomparisoninterpreterresult} will now be normalized. 
Since weights of -10 to 10 are possible, there are 21 values, which are scaled from 0 to 20.
Those will then be multiplied with the implementation score and then accumulated.
Since there are 11 requirements which could be implemented with a score of 4 and a median weight of 10, the \textbf{maximum possible score is 440}.

\begin{table}[h]
	\begin{center}
		\begin{tabular}{r|l|l|l|l}
			& Requirement       & Weight & Quality & Score \\ \hline
			F1 & Reflects changes to the model    & 16 & 4 & 64    \\
			NF2 & Fast app start    & 14 & 4 & 56   \\
			NF1 & Low battery usage            &16  & 3  & 48   \\
			F5 & Poor configurations lead to fall-backs      & 16 & 3 & 48    \\
			NF5 & Small download size    & 12 & 3 & 36    \\
			NF3 & High UI and animation performance      & 14    & 2  & 28 \\
			F2 & Offline capable       & 8 & 3 & 24   \\
			NF4 & Low network usage      & 4    & 4  & 16 \\
			F3 & Cache-first policy       & 4  & 3 & 12  \\
			NF6 & Small app size on disk      & 4  & 3 & 12  \\
			F4 & Clear cache and data       & 2 & 3   & 6   
			\\ \hline
			$\sum$ &    & &    &   \textbf{350}
		\end{tabular}
	\end{center}
	\caption{Implementation score of each requirement of the native approach interpreter.}
	\label{table:implementationscorenativeapproachresults}
\end{table}

Table \ref{table:implementationscorenativeapproachresults} shows, that the end score of the native approach interpreter is 350 of 440.
The table also shows, that the requirement NF3 (High UI and animation performance) is of above-average importance (14 out of 20), but the implementation quality is only scored at 2 out of 4.
This represents the main issue with the native approach interpreter implementation.

\subsection{Virtual Machine Approach}

Since there is no full implementation of a virtual machine approach, only assumptions can be made.
Since the proof-of-concept implementation implies a buildable architecture, the following table (\ref{table:implementationscorevirtualmachineapproachresults}) assumes, that there is a correct and full implementation of the virtual machine approach.

\begin{table}[h]
	\begin{center}
		\begin{tabular}{r|l|l|l|l}
			& Requirement       & Weight & Imaginary Quality & Score \\ \hline
			F1 & Reflects changes to the model    & 16 & 4 & 64    \\
			NF2 & Fast app start    & 14 & 4 & 56   \\
			NF3 & High UI and animation performance      & 14    & 4  & 56 \\
			NF1 & Low battery usage            &16  & 3  & 48   \\
			F5 & Poor configurations lead to fall-backs      & 16 & 3 & 48    \\
			NF5 & Small download size    & 12 & 3 & 36    \\
			F2 & Offline capable       & 8 & 3 & 24   \\
			NF4 & Low network usage      & 4    & 4  & 16 \\
			F3 & Cache-first policy       & 4  & 3 & 12  \\
			NF6 & Small app size on disk      & 4  & 3 & 12  \\
			F4 & Clear cache and data       & 2 & 3   & 6   
			\\ \hline
			$\sum$ &    & &    &   \textbf{378}
		\end{tabular}
	\end{center}
	\caption{Implementation score of an imaginary virtual machine implementation.}
	\label{table:implementationscorevirtualmachineapproachresults}
\end{table}

It can be assumed that the other requirements are not affected.
This way, the score increases from 350 to 378 (out of 440).
This is an increase from 79.55 \% to 85.91 \% which is an improvement of 6.36 \%.

\section{Threats to Validity}

The scores that have been evaluated in the previous section are subject to a possible invalidity, because there are internal issues with the process.
Since this evaluation is not a study but a scoring procedure, the common threats to validity cannot be applied directly.
That is why new and suitable threats will be defined and discussed.
The following threats refer to the internal validity of the scoring process and result.
There is no external validity, because the result only indicates the concrete implementation quality for the given set of requirements.

\begin{enumerate}
	\item \textbf{Not enough requirements are used in the process.}
	
	If there are not enough requirements used in the process the remaining ones are all more relevant and slight changes in the weight classification have a strong impact on the result.
	The worst case scenario is that the most important requirements are missing, so the less relevant ones have the most significant impact on the score, which will result in wrong assumptions.
	A comparison of two implementations is then only valid for exactly those requirements.
	
	\item \textbf{The weights were evaluated incorrectly.}
	
	Wrong weights could be the result of misinterpreted requirements, because they either are not sufficiently specified or their relevance was wrongly assessed.
	To address this, the process should be done in a Team of experts and product owners, so that functional and non-functional requirements can be evaluated.
	A product owner might rate the non-functional requirements like download size and battery consumption much higher than the development team.
	
	\item \textbf{The implementation quality scores were evaluated incorrectly.}
	
	This might happen if the actual quality is unknown, because it was not sufficiently evaluated beforehand.
	Another common mistake is, that instead of evaluating the quality of the implementation, the quality or correctness of the requirement is scored.
	An example is the requirement \textit{F1 Reflects changes to the model}.
	While the native approach implementation perfectly implements the required functionality, the requirement itself is insufficient and must be extended.
	An example are cached images that have been updated on the server.
	They will not be invalidated and updated, because the \ac{URL} did not change and images will be cached infinitely.
	But since the interpreter implements the required functionality, the implementation quality is full (4 out of 4).
	
	\item \textbf{The score is misinterpreted as a general software quality.}
	
	The end score is only meaningful with regard to the maximum score and the score of alternative implementations.
	Also, the score is only telling the team how well the application's requirements are implemented.
	That refers directly to the statements of \cite{klein_evaluating_2001} which describe that a software system can only be evaluated against a set of requirements.
	There is no general quality statement.
	
	\item \textbf{Vague assumptions caused by minor score differences}
	
	In this example, the second approach has only a 6.36 \% higher score, which is only due to one single requirement.
	This does not define a clear winner, because the various scores are to similar.
	This is why, it is not possible to conclusively state that the virtual machine approach is significantly better than the current implementation.
	
\end{enumerate}









